A Partitioning Algorithm for Maximum Common Subgraph Problems\thanks{This work was supported by a pile of money from a funding agency.}

### 2017-02-16 paper/paper.tex
Bron Kerbosch citation
Muck around with citations

To determine the similarity or difference between two graphs, we must first
find what they have in common
\citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,KriegeThesis}.
The \emph{maximum common subgraph} family of problems involves finding a large
graph which is isomorphic to subgraphs of two given graphs simultaneously.
Because graphs are widely used to model real-world phenomena, maximum common
subgraph problems have arisen in molecular science (where graphs represent
molecules)
\citep{DBLP:journals/jcamd/RaymondW02a,Ehrlich:2011,DAM2014,Grindley1993707},
computer vision \cite{DBLP:journals/jair/CookH94}, discovery of repeated
patterns in source code \cite{DBLP:journals/tkde/DjokoCH97} and circuits
\cite{DBLP:journals/jair/CookH94}, and recognition of handwritten characters
\cite{SIWEILU1991617}.

Maximum common subgraph problems are NP-hard, and remain challenging
computationally. Recent practical progress has been made by using constraint
programming \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16} and
mathematical programming \citep{DBLP:journals/dam/BahienseMPS12}, by reducing
to the maximum clique problem \citep{LeviG,DBLP:conf/cp/McCreeshNPS16}, and by
adapting subgraph isomorphism algorithms \citep{UpcomingAAAIPaper}. This paper
introduces a new branch and bound algorithm which exploits special properties
of these problems to allow a much faster exploration of the search space,
whilst retaining the filtering and bounding benefits of the constraint
programming approach. We describe the algorithm for the basic maximum common
subgraph problem, and discuss how it may be adapted to handle vertex labels,
edge labels, and the requirement that the found subgraph be connected. We then
present an empirical study of the algorithm, demonstrating that it improves the
state of the art by more than an order of magnitude on the unlabelled variant
of the problem, and showing that it can handle much larger instances than
earlier constraint programming or clique approaches due to lower memory usage.

@@

\paragraph{Label classes} We require only $\BigO{g+h}$ space per level of the
search tree to store labelling information.  This is done by storing a
\emph{label class} as a pair $\langle P,T \rangle$ for each label $l$ that is
used, where $P$ is the set of vertices in $\V(G)$ labelled $l$, and $T$ is the
set of vertices in $\V(H)$ labelled $l$\footnote{We adopt terminology from
subgraph isomorphism, where one graph is a pattern graph $P$ and the other the
target graph $T$.}. Since there are $g + h$ vertices in the two graphs, at most
$g + h$ label classes can exist at once, and there are at most $g + h$ vertices
in the union of all of the $P$ and $T$ sets. Furthermore, we do not actually
need to store the bits making up a label---we care only that like-labelled
vertices are kept together, and the label itself is not used. Nor do we need to
store any label class which is present only in one graph but not the other (or
which is not present at all).  Together, these facts allow us to store all the
necessary information in two pairs of flat arrays. One pair of arrays is used
to store a permutation of the vertices in $P$ having like-labelled vertices
stored consecutively, together with the indices of the start of each label in
turn (in some arbitrary order). The second pair stores the permutation of
vertices in $T$, together with label start indices \emph{in the same order} as
they are used in the first pair of arrays. ?? This representation has
similarities to data structures used in partition backtracking and bron
kerbosch.

>>>>

To determine the similarity or difference between two graphs, we must first
find what they have in common
\citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,KriegeThesis}.
The \emph{maximum common subgraph} family of problems involves finding a large
graph which is isomorphic to subgraphs of two given graphs simultaneously.
Because graphs are widely used to model real-world phenomena, maximum common
subgraph problems have arisen in molecular science (where graphs represent
molecules)
\citep{DBLP:journals/jcamd/RaymondW02a,Ehrlich:2011,DAM2014,Grindley1993707},
and also in other domains including malware detection
\citep{DBLP:journals/compsec/ParkRS13}, source code analysis
\cite{DBLP:journals/tkde/DjokoCH97}, and computer vision
\cite{DBLP:journals/jair/CookH94}.

Maximum common subgraph problems are NP-hard, and remain challenging
computationally. Recent practical progress has been made by using constraint
programming \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16} and
mathematical programming \citep{DBLP:journals/dam/BahienseMPS12}, by reducing
to the maximum clique problem \citep{LeviG,DBLP:conf/cp/McCreeshNPS16}, and by
adapting subgraph isomorphism algorithms \citep{UpcomingAAAIPaper}. Some
special cases also have practial polynomial time algorithms
\citep{DBLP:conf/mfcs/DroschinskyKM16,DBLP:conf/sofsem/DroschinskyKM17}.

This paper introduces a new branch and bound algorithm which exploits special
properties of these problems to allow a much faster exploration of the search
space, whilst retaining the filtering and bounding benefits of the constraint
programming approach. We describe the algorithm for the basic maximum common
subgraph problem, and discuss how it may be adapted to handle vertex labels,
edge labels, and the requirement that the found subgraph be connected. We then
present an empirical study of the algorithm, demonstrating that it improves the
state of the art by more than an order of magnitude on the unlabelled variant
of the problem, and showing that it can handle much larger instances than
earlier constraint programming or clique approaches due to lower memory usage.

@@

\paragraph{Label classes} We require only $\BigO{g+h}$ space per level of the
search tree to store labelling information.  This is done by storing a
\emph{label class} as a pair $\langle P,T \rangle$ for each label $l$ that is
used, where $P$ is the set of vertices in $\V(G)$ labelled $l$, and $T$ is the
set of vertices in $\V(H)$ labelled $l$\footnote{We adopt terminology from
subgraph isomorphism, where one graph is a pattern graph $P$ and the other the
target graph $T$.}. Since there are $g + h$ vertices in the two graphs, at most
$g + h$ label classes can exist at once, and there are at most $g + h$ vertices
in the union of all of the $P$ and $T$ sets. Furthermore, we do not actually
need to store the bits making up a label---we care only that like-labelled
vertices are kept together, and the label itself is not used. Nor do we need to
store any label class which is present only in one graph but not the other (or
which is not present at all).  Together, these facts allow us to store all the
necessary information in two pairs of flat arrays. One pair of arrays is used
to store a permutation of the vertices in $P$ having like-labelled vertices
stored consecutively, together with the indices of the start of each label in
turn (in some arbitrary order). The second pair stores the permutation of
vertices in $T$, together with label start indices \emph{in the same order} as
they are used in the first pair of arrays. ?? This representation has
similarities to data structures used in partition backtracking and the \citet{DBLP:journals/cacm/BronK73} clique enumeration algorithm.
